---
title: "Confirmatory Analysis"
author: "Justin Lim Zheng Han"
date: "March 19, 2025"
date-modified: "last-modified" 
---

# Confirmatory Data Analysis

## Getting Started

### Installing the required R packages

We will load the following packages by using the pacman::p_load function:

tidyquant, tidyverse, xts, timeDate, lubridate, writexl, broom, sandwich, lmtest, dplyr, tidyr, GGally, corrplot, zoo, plotly, scales, moments

```{r}
#| echo: false

pacman::p_load(tidyquant, tidyverse, xts, timeDate, lubridate, writexl, broom, sandwich, lmtest, dplyr, tidyr, GGally, corrplot, zoo, plotly, scales, moments)
```

### Loading the Dataset

Our stock data will be retrieved from Yahoo Finance using quantmod. In Shiny App, Data will be pulled real-time according to user's selection. For now below data will be used to demonstrate EDA and CDA.

### 📦 Simulated User Portfolio Input

```{r}
# Simulate user portfolio (can be dynamic in Shiny later) 
user_portfolio <- data.frame(
  symbol = c("AAPL", "MSFT", "TSLA", "NVDA", "AMZN", "NVDA"),
  buy_date = as.Date(c("2022-01-15", "2022-02-20", "2022-03-10", "2022-04-05", "2022-05-12", "2024-01-17")),
  quantity = c(10, 15, 5, 8, 12, 7)
)
user_portfolio
```

## Data Overview (Exploratory Data Analysis)

### Portfolio-Level Summary

In R Shiny, user will be able to select the stocks they have in their Portfolio. From the selected stocks, EDA will run accordingly.

### 📈 Download Stock Price Data from Yahoo Finance

```{r}
# Get stock data from buy dates onward
data_start <- min(user_portfolio$buy_date)
stock_data <- tq_get(user_portfolio$symbol, from = data_start, to = Sys.Date())
```

### 📊 Calculate Returns and Current Values

```{r}
# Join each stock with its buy date and quantity
portfolio_results <- list()

for (i in 1:nrow(user_portfolio)) {
  stock <- user_portfolio$symbol[i]
  buy_date <- user_portfolio$buy_date[i]
  qty <- user_portfolio$quantity[i]

  # Get prices for that stock
  stock_prices <- stock_data %>% filter(symbol == stock)

  if (nrow(stock_prices) == 0) {
    warning(paste("No price data found for", stock, "- skipping..."))
    next
  }

  available_dates <- stock_prices$date
  valid_buy_date <- max(available_dates[available_dates <= buy_date])

  # Check if a valid buy date was found
  if (is.infinite(valid_buy_date) | is.na(valid_buy_date)) {
    warning(paste("No available trading date before", buy_date, "for", stock, "- skipping..."))
    next
  }

  # Get the adjusted price on the valid buy date
  buy_price_row <- stock_prices %>% filter(date == valid_buy_date)
  if (nrow(buy_price_row) == 0) {
    warning(paste("Buy price not available on", valid_buy_date, "for", stock, "- skipping..."))
    next
  }

  buy_price <- buy_price_row$adjusted[1]

  # Proceed only if buy_price is non-missing and > 0
  if (is.na(buy_price) || buy_price == 0) {
    warning(paste("Invalid buy price for", stock, "- skipping..."))
    next
  }

  # Build return tracking table
  tracking_data <- stock_prices %>%
    filter(date >= valid_buy_date) %>%
    arrange(date) %>%
    mutate(
      return_since_buy = (adjusted / buy_price) - 1,
      current_value = adjusted * qty,
      stock = stock,
      buy_date = valid_buy_date,
      quantity = qty,
      buy_price = buy_price,
      invested_amount = buy_price * qty,
      holding_value = adjusted * qty,
      unrealized_return = (adjusted - buy_price) * qty
    )

  portfolio_results[[i]] <- tracking_data
}

portfolio_tracking_all <- bind_rows(portfolio_results)
```

### 📋 Portfolio Summary (Latest Value & Return)

```{r}
# Add position_id to tracking data
portfolio_tracking_all <- portfolio_tracking_all %>%
  mutate(position_id = paste0(stock, "_", buy_date))

# Group by position_id (instead of just stock)
portfolio_summary <- portfolio_tracking_all %>%
  group_by(position_id) %>%
  filter(date == max(date)) %>%
  summarise(
    stock = first(stock),
    buy_date = first(buy_date),
    quantity = first(quantity),
    buy_price = first(buy_price),
    invested_amount = first(invested_amount),
    current_price = first(adjusted),
    current_value = first(holding_value),
    unrealized_return = first(unrealized_return),
    return_percent = first(return_since_buy)
  ) %>%
  mutate(return_percent_label = scales::percent(return_percent))

portfolio_summary
```

### 📈 Plot Return Since Buy Date (All Stocks)

```{r}
# Base ggplot chart with distinct lines for same stock different buy-ins
p <- ggplot(portfolio_tracking_all, aes(x = date, y = return_since_buy, color = position_id)) +
  geom_line(size = 1) +
  labs(title = "Cumulative Return Since Buy Date (By Position)", y = "Return", x = "Date") +
  scale_y_continuous(labels = scales::percent)

# Make it interactive
ggplotly(p, tooltip = c("x", "y", "color"))
```

### 📈 Interactive Plot Return for each Stock

```{r}
plot_ly(
  data = portfolio_tracking_all,
  x = ~date,
  y = ~holding_value,
  color = ~position_id,
  type = 'scatter',
  mode = 'lines',
  text = ~paste0(
    "Stock: ", stock, "<br>",
    "Buy Date: ", buy_date, "<br>",
    "Date: ", format(date, "%Y-%m-%d"), "<br>",
    "Value: $", format(round(holding_value, 2), big.mark = ",")
  ),
  hoverinfo = "text"
) %>%
  layout(
    title = "Individual Stock Value Over Time (By Buy-in Position)",
    xaxis = list(title = "Date"),
    yaxis = list(title = "Holding Value (USD)")
  )
```

### 📦 Total Portfolio Value Over Time

```{r}
total_portfolio <- portfolio_tracking_all %>%
  group_by(date) %>%
  summarise(total_value = sum(holding_value), .groups = "drop")

# ggplot base
p <- ggplot(total_portfolio, aes(x = date, y = total_value)) +
  geom_line(color = "darkgreen", size = 1.2) +
  labs(title = "Total Portfolio Value Over Time", y = "Value (USD)", x = "Date") +
  scale_y_continuous(labels = scales::dollar)

# Convert to interactive
ggplotly(p, tooltip = c("x", "y"))
```

### ✅ Summary

-   Simulated multi-stock portfolio
-   Live Yahoo Finance data via `tq_get()`
-   Return tracking since individual buy dates
-   Portfolio-level return and value plots

➡ Ready to convert into Shiny later using `selectInput()` and `dateInput()`!

```{r}
colnames(portfolio_tracking_all)
```

#### Time-series Price Trend

```{r}
ggplot(stock_data, aes(x = date, y = adjusted, color = symbol)) +
  geom_line() +
  labs(title = "Adjusted Closing Price Over Time", y = "Adjusted Price", x = "Date")
```

#### Rolling

```{r}
stock_data <- stock_data %>%
  group_by(symbol) %>%
  arrange(date) %>%
  mutate(daily_return = adjusted / lag(adjusted) - 1) %>%
  ungroup()

stock_data <- stock_data %>%
  group_by(symbol) %>%
  arrange(date) %>%
  mutate(rolling_vol_20 = rollapply(daily_return, width = 20, FUN = sd, fill = NA, align = "right")) %>%
  ungroup()

ggplot(stock_data %>% filter(!is.na(rolling_vol_20)), 
       aes(x = date, y = rolling_vol_20, color = symbol)) +
  geom_line(size = 1) +
  labs(
    title = "20-Day Rolling Volatility (Standard Deviation of Daily Returns)",
    x = "Date",
    y = "Rolling Volatility (Std Dev)"
  ) +
  scale_y_continuous(labels = scales::percent) +
  theme_minimal()
```

Observations:

-   TSLA (Purple Line) shows the highest volatility overall. It fluctuates often and peaks above 6%, meaning high price swings daily.

-   MSFT (Green Line) is the least volatile, maintaining a very consistent low volatility around 1–2%, reflecting its status as a stable, mature stock.

-   NVDA (Cyan) and AMZN (Red) fluctuate in between — but note that NVDA has occasional sharp volatility spikes, likely due to earnings or news events.

-   Spikes in volatility (upward jumps) are usually tied to market news, economic events, earnings releases, or crashes/rallies.

-   Periods of low volatility indicate price stability or consolidation phases.

### Individual stock-level EDA

#### Descriptive Statistics

```{r}
portfolio_tracking_all %>%
  group_by(position_id) %>%
  summarise(
    stock = first(stock),
    buy_date = first(buy_date),
    mean_return = mean(return_since_buy, na.rm = TRUE),
    sd_return = sd(return_since_buy, na.rm = TRUE),
    median_return = median(return_since_buy, na.rm = TRUE),
    min_return = min(return_since_buy, na.rm = TRUE),
    max_return = max(return_since_buy, na.rm = TRUE)
  )
```

mean_return: Daily gain on average sd_return: Daily volatility median_return: Median min_return: Minimum return max_return: Maximum return

#### Cumulative Returns

```{r}
portfolio_tracking_all <- portfolio_tracking_all %>%
  arrange(stock, buy_date, date) %>%
  group_by(position_id) %>%
  mutate(
    cumulative_return = cumprod(1 + coalesce(return_since_buy, 0))
  ) %>%
  ungroup()

ggplot(portfolio_tracking_all, aes(x = date, y = return_since_buy, color = position_id)) +
  geom_line(size = 1) +
  labs(
    title = "Cumulative Return Since Buy Date (Per Buy-in)",
    x = "Date",
    y = "Cumulative Return"
  ) +
  scale_y_continuous(labels = scales::percent) +
  theme_minimal()
```

### Correlation analysis

```{r}
# Ensure daily return is available (if not already calculated)
stock_data <- stock_data %>%
  group_by(symbol) %>%
  arrange(date) %>%
  mutate(daily_return = adjusted / lag(adjusted) - 1) %>%
  ungroup()

# Summarise to ensure one return per symbol-date (safety step)
returns_clean <- stock_data %>%
  group_by(date, symbol) %>%
  summarise(daily_return = mean(daily_return, na.rm = TRUE), .groups = "drop")

# Pivot wider
returns_matrix <- returns_clean %>%
  pivot_wider(names_from = symbol, values_from = daily_return)

# Ensure numeric
returns_matrix <- returns_matrix %>%
  mutate(across(-date, as.numeric))

# Compute correlation matrix
returns_corr <- cor(returns_matrix %>% select(-date), use = "pairwise.complete.obs")

# Plot
corrplot(
  returns_corr,
  method = "color",
  addCoef.col = "black",
  number.cex = 0.8,
  title = "Daily Return Correlation Matrix",
  mar = c(0, 0, 2, 0)
)
```

Observation:

-   AMZN and MSFT show the highest correlation in your portfolio — likely due to their similar sector dynamics (large-cap tech, index-driven movements).

#### Return Vs Votality

```{r}
# Step 1: Calculate mean return and volatility per stock
return_volatility_stats <- stock_data %>%
  group_by(symbol) %>%
  summarise(
    mean_return = mean(daily_return, na.rm = TRUE),
    volatility = sd(daily_return, na.rm = TRUE),
    .groups = "drop"
  )

# Step 2: Plot the Risk–Return scatter plot
ggplot(return_volatility_stats, aes(x = volatility, y = mean_return, label = symbol)) +
  geom_point(size = 4, color = "darkblue") +
  geom_text(nudge_y = 0.0005, fontface = "bold") +
  labs(
    title = "Return vs Volatility (Risk–Return Tradeoff)",
    x = "Volatility (Standard Deviation of Daily Return)",
    y = "Mean Daily Return"
  ) +
  scale_x_continuous(labels = scales::percent) +
  scale_y_continuous(labels = scales::percent) +
  theme_minimal()
```

## Data Overview (CDA)

### 1. Wilcoxon Rank-Sum Test (Median Comparison)

```{r}
# Replace with user-defined stock selection
stock_a <- "AMZN"
stock_b <- "TSLA"

df_wilcox <- stock_data %>%
  filter(symbol %in% c(stock_a, stock_b)) %>%
  select(symbol, daily_return) %>%
  filter(!is.na(daily_return))

wilcox.test(daily_return ~ symbol, data = df_wilcox)
```

Observation:

-   p-value = 0.6222 is much higher than 0.05 → So you fail to reject the null hypothesis.

-   There is no statistically significant difference in the median daily returns of AMZN and TSLA.

### 2. F-Test for Variance Comparison

```{r}
# 📌 Simulated user input (replace with input$stock1 / input$stock2 in Shiny later)
stock1 <- "AMZN"
stock2 <- "MSFT"

# Step 1: Prepare data for selected stocks
var_test_data <- stock_data %>%
  filter(symbol %in% c(stock1, stock2)) %>%
  select(symbol, daily_return) %>%
  filter(!is.na(daily_return))

# Step 2: Perform F-test
cat(paste0("Comparing Variance of Daily Returns: ", stock1, " vs ", stock2, "\n"))
var_test_result <- var.test(daily_return ~ symbol, data = var_test_data)
print(var_test_result)
```

### 3. Pearson Correlation Test (Relationship)

```{r}
stock1 <- "AMZN"
stock2 <- "MSFT"

cor_data <- stock_data %>%
  filter(symbol %in% c(stock1, stock2)) %>%
  select(symbol, date, daily_return) %>%
  pivot_wider(names_from = symbol, values_from = daily_return) %>%
  filter(!is.na(.data[[stock1]]), !is.na(.data[[stock2]]))

cor.test(cor_data[[stock1]], cor_data[[stock2]], method = "pearson")
```

### 4. Multiple Linear Regression: Return \~ Volume + Market Index

```{r}
# --- 📌 Simulated user inputs (can be replaced by Shiny inputs) ---
target_stock <- "AMZN"
market_index_stock <- "MSFT"

# --- Step 1: Prepare dataset for regression ---
reg_data <- stock_data %>%
  filter(symbol %in% c(target_stock, market_index_stock)) %>%
  select(date, symbol, daily_return, volume) %>%
  pivot_wider(names_from = symbol, values_from = c(daily_return, volume))

# --- Step 2: Rename columns dynamically ---
# Construct variable names
target_return_col <- paste0("daily_return_", target_stock)
target_volume_col <- paste0("volume_", target_stock)
market_return_col <- paste0("daily_return_", market_index_stock)

# Rename for regression model
reg_data <- reg_data %>%
  rename(
    target_return = !!target_return_col,
    target_volume = !!target_volume_col,
    market_return = !!market_return_col
  ) %>%
  filter(!is.na(target_return), !is.na(target_volume), !is.na(market_return))

# --- Step 3: Run Multiple Linear Regression ---
cat(paste0("📊 Multiple Linear Regression: ", target_stock, " ~ Volume + ", market_index_stock, " Return\n"))
model <- lm(target_return ~ target_volume + market_return, data = reg_data)
summary(model)
```

### 5. ANOVA Test (One-way or Two-way)

```{r}
# One-way ANOVA: Do average daily returns differ across your stocks?
anova_data <- stock_data %>%
  select(symbol, daily_return) %>%
  filter(!is.na(daily_return))

anova_model <- aov(daily_return ~ symbol, data = anova_data)

# Display ANOVA summary
cat("📊 One-Way ANOVA: Daily Return ~ Stock\n")
summary(anova_model)
```

### 5. Two-way ANOVA (Return \~ Symbol + Month)

```{r}
# Add a "month" variable
stock_data <- stock_data %>%
  mutate(month = month(date, label = TRUE))

# Prepare data
anova2_data <- stock_data %>%
  select(symbol, daily_return, month) %>%
  filter(!is.na(daily_return))

# Two-way ANOVA: Test effect of both stock and month
anova2_model <- aov(daily_return ~ symbol + month, data = anova2_data)

# Display ANOVA summary
cat("📊 Two-Way ANOVA: Daily Return ~ Stock + Month\n")
summary(anova2_model)
```

### 6. Event Study Analysis – Cumulative Abnormal Returns (CAR) during COVID (2020–2022)

```{r}
# --- 📌 Simulated user input (replace with selectInput in Shiny later) ---
target_stock <- "NVDA"
market_index_stock <- "MSFT"

# Step 1: Get event period stock data (COVID years 2020–2022)
event_stocks <- unique(c(target_stock, market_index_stock))

event_stock_data <- tq_get(event_stocks, from = "2020-01-01", to = "2022-12-31") %>%
  group_by(symbol) %>%
  arrange(date) %>%
  mutate(daily_return = adjusted / lag(adjusted) - 1) %>%
  ungroup()

# Step 2: Pivot wider for return matrix
event_df <- event_stock_data %>%
  select(date, symbol, daily_return) %>%
  filter(!is.na(daily_return)) %>%
  pivot_wider(names_from = symbol, values_from = daily_return)

# Step 3: Rename columns dynamically
target_col <- target_stock
market_col <- market_index_stock
colnames(event_df)[which(colnames(event_df) == target_stock)] <- "target_return"
colnames(event_df)[which(colnames(event_df) == market_index_stock)] <- "market_return"

# Step 4: Estimate expected returns (OLS regression)
model <- lm(target_return ~ market_return, data = event_df)

# Step 5: Compute abnormal return & CAR
event_df$expected_return <- predict(model, newdata = event_df)
event_df <- event_df %>%
  mutate(
    abnormal_return = target_return - expected_return,
    CAR = cumsum(abnormal_return)
  )

# Step 6: Plot CAR interactively
plot_ly(event_df, x = ~date, y = ~CAR, type = "scatter", mode = "lines",
        line = list(color = "darkblue")) %>%
  layout(
    title = paste(target_stock, "Cumulative Abnormal Return (2020–2022)"),
    xaxis = list(title = "Date"),
    yaxis = list(title = "Cumulative Abnormal Return (CAR)")
  )
```
